Planning to Save Energy: How Information Format Affects Accuracy

### 2024-12-04 manuscript/manuscript.qmd
introduction outline. Fixed plot bug.

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
author:
    - name: Thomas E. Gorman
      url: https://tegorman13.github.io/
      orcid: "0000-0001-5366-5442"
      affiliations: 
        - name:  Purdue University, USA
          city: West Lafayette
          region: IN
          country: USA
          postal-code: 47907
          affiliation-url: https://web.ics.purdue.edu/~treimer/
      attributes: 
          corresponding: true
    - name: Torsten Reimer
      affiliation: Brian Lamb School of Communication, Purdue University
      affiliation-url: https://cla.purdue.edu/communication/
      url: https://web.ics.purdue.edu/~treimer/
      orcid: 0000-0002-7419-0076
    - name: Juan Pablo Loaiza Ramirez
      affiliation: Brian Lamb School of Communication, Purdue University
      affiliation-url: https://cla.purdue.edu/communication/
      url: https://web.ics.purdue.edu/~treimer/
      orcid: 0000-0001-9663-0522
format:
    # html: 
    #     grid:
    #         sidebar-width: 220px
    #         body-width: 1200px
    #         margin-width: 170px
    #         gutter-width: 1.0rem
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        code-fold: true
        code-summary: "Show the code"
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    # hugo-md:
    #     include: true
    #     html-math-method: mathjax
    #     output-file: manuscript_hugo.md
    gfm:
        echo: true
        output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
# format-links:   
#     - html
#     - gfm
#     - format: hikmah-manuscript-pdf
#       text: PDF version
#       icon: file-pdf
date: 2024-11-19
abstract: |
    This study examines how presenting energy information in different formats (kWh, %, USD) affects planning accuracy. Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation.
---

@@

@@

@@

@@

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
    # scale_x_continuous("y", breaks = 1:7) +
    # scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
    ggtitle("Data with posterior predictions",
            subtitle = "S1. red bars are empirical data, black points are model predictions and credible intervals") +
    labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
    theme_minimal()
    theme(legend.background = element_blank() #,legend.position = c(.9, .75)
    )

@@

s2_agg1 |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
    filter(plan=="plan1",rounded=="Rounded") |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  filter(mean_abs_error <= 0.50) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7))) + 
    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
                    labels = scales::percent_format(accuracy = 1))

>>>>

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
author:
  - name: Thomas E. Gorman
    url: https://tegorman13.github.io/
    orcid: "0000-0001-5366-5442"
    affiliations: 
      - ref: cclab
      - ref: clara
    attributes: 
      corresponding: true
  - name: Torsten Reimer
    url: https://web.ics.purdue.edu/~treimer/
    orcid: 0000-0002-7419-0076
    affiliations:
      - ref: cclab
      - ref: clara
  - name: Juan Pablo Loaiza Ramirez
    url: https://www.cla.purdue.edu/directory/profiles/loaiza-ramirez,juan-pablo.html
    orcid: 0000-0001-9663-0522
    affiliations:
      - ref: cclab
      - ref: clara
  - name: Hayden Barber
    url: https://www.sdstate.edu/directory/hayden-barber
    orcid: 0000-0002-3465-8615
    affiliations:
      - ref: sdsu
affiliations:
  - id: cclab
    name: Communication and Cognition Lab, Purdue University
    city: West Lafayette
    region: IN
    country: USA
    postal-code: 47907
    url: https://web.ics.purdue.edu/~treimer/
  - id: clara
    name: College of Liberal Arts Research Academy
    department: Purdue University
    city: West Lafayette
    region: IN
    country: USA
    postal-code: 47907
    url: https://cla.purdue.edu/about/college-initiatives/research-academy/
  - id: sdsu
    name: School of Communication & Journalism, South Dakota State University
    city: Brookings
    region: SD
    country: USA
    url: https://www.sdstate.edu/directory/hayden-barber
format:
    # html: 
    #     grid:
    #         sidebar-width: 220px
    #         body-width: 1200px
    #         margin-width: 170px
    #         gutter-width: 1.0rem
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        code-fold: true
        code-summary: "Show the code"
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    hugo-md:
        include: true
        echo: false
        html-math-method: mathjax
        output-file: manuscript_hugo.md
    gfm:
        echo: true
        output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    hikmah-manuscript-docx: default
# format-links:   
#     - html
#     - gfm
#     - format: hikmah-manuscript-pdf
#       text: PDF version
#       icon: file-pdf
date: 2024-12-03
abstract: |
    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans. By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
---

@@

- highlight the significant contribution of residential energy consumption to carbon emissions and the potential for substantial reductions. Make point about urgency of climate change? (IPCC (2014) and EIA (2012) ).

- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption (Fischer, 2008)

@@

- Canfield et al. (2017) found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading. However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
- Canfield et al.'s (2017) findings on preferences for historical use information and the impact of neighbor comparisons
- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements (Vessey, 1991)
- alignment can reduce cognitive load and enhance accuracy in planning (Shah & Freedman, 2009)
- Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.

@@

The way numerical information is presented can significantly affect how individuals process and use that information (Reimer et al., 2015). The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to (Gigerenzer & Edwards, 2003; Reimer et al., 2015). Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.

Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities (Gigerenzer & Hoffrage, 1995; Hoffrage et al., 2000). In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.

Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers’ ability to create accurate energy conservation plans. Specifically, there is a gap in understanding how presenting energy information in absolute units versus percentages or monetary terms affects the precision of planning appliance-specific reductions. Addressing this gap is crucial for developing effective interventions that promote energy conservation behaviors.

@@

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
    # scale_x_continuous("y", breaks = 1:7) +
    # scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
    ggtitle("Data with posterior predictions",
            subtitle = "S1. red bars are empirical data, black points are model predictions and credible intervals") +
    labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
    theme_minimal() +
    theme(legend.background = element_blank() #,legend.position = c(.9, .75)
    )

@@

s2_agg1 |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
    filter(plan=="plan1",rounded=="Rounded") |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  filter(mean_abs_error <= 0.50) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y=element_text(face = "plain", size = rel(0.7))) + 
    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
                    labels = scales::percent_format(accuracy = 1))
